---
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)


```

## 3a)

The response quantitative variable is the number of bacteria present on the second recipient gloved hand.

the predictor variable includes a single factor that is the methode of greeting with the levels/treatment being a high five and handshake. This is a categorical variable. 


## 3b)


```{r}

#data
Handshake <- c(115, 87, 109, 96, 92, 90, 102, 110, 89, 100)
HighFive <- c(48, 70, 69, 60, 53, 56, 80, 74, 48, 55)

 
treatments<- c("Handshake", "HighFive")
boxplot(Handshake, HighFive, names=treatments,  main = "Bacteria count boxplot", 
xlab="Method of greeting", ylab="Bacteria count")

```


It is believed that more bacteria were transferred during a handshake compared with a high five.
This is supported by the box plots as we can see, the entire boxplot of handshake is above the entire box plot for the high five.
This means that the min, q1, median, q3 and max values of the number of bacteria in handshakes are all higher than the max value for a high five.
So we can conclude that the boxplots do support the original theory. 


## 3c)


```{r}


print("Handshake")
qqnorm(Handshake)
qqline(Handshake)
print("HighFive")
qqnorm(HighFive)
qqline(HighFive)

a<-var.test(Handshake,HighFive)
a


b<-t.test(Handshake,HighFive,var.equal=TRUE)
b


```

assumptions

normality: Using the qq plots for handshakes and high fives we see that both are sets of data follow the assumptions of normality. 

variance: Using var.test the value came out to 0.6955 which is a lot higher than 0.05. From this value we can tell the variances are roughly equal and the assumption is followed.

Both normality and variance assumption were followed. 

using t.test welsh approximation the p value was 2.534e-07 (much lower than 0.05) which means we can reject the null hypothesis and this information in conjunction with the boxplot lets us conclude that more bacteria were transferred during a handshake compared with a high five.